Redshift Evolution of the Underlying Type Ia Supernova Stretch Distribution

### 2020-03-26 Nicolas.tex
Greg corr

\begin{document}
\title{Redshift Evolution of the Type Ia Supernova Population Stretch
Distribution}

\author{N. Nicolas
\thanks{n.nicolas@ipnl.in2p3.fr, equal contribution}
\inst{1}
    \and M. Rigault
\thanks{m.rigault@ipnl.in2p3.fr, equal contribution}
\inst{2}
    \and R. Graziani\inst{2}
    \and M. Briday\inst{1}
    \and Y. Copin\inst{1}
    \and Y.-L. Kim\inst{1}
}

\institute{Université de Lyon, F-69622, Lyon, France; Université de Lyon
    1, Villeurbanne; CNRS/IN2P3, Institut de Physique des Deux Infinis, Lyon
    \and Université Clermont Auvergne, CNRS/IN2P3, Laboratoire de
    Physique de Clermont, F-63000 Clermont-Ferrand, France.
}

@@

\abstract{\mr{ABSTRACT NOT DONE YET.}
Type Ia supernovae (SNe Ia) allow for the construction of the Hubble
    diagram, giving us information about the Universe's expansion and its
    fundamental components, one of which is dark energy. But systematic
    uncertainties are now starting to be limiting in our ability to measure
    those parameters. In particular, the physics of SNe Ia is still mostly
    unknown, and is thought not to change in time/with the redshift.}
    {In an attempt to reduce those uncertainties, we try to find an empirical
    law describing SNe Ia's length of explosion (stretch) evolution with the
redshift.}
    {We started by getting a complete sample representing all of the stretch
        distribution that Nature can give us, before using LsSFR measurements,
        an age tracer which evolution with redshift is known, that has been
        shown to have a strong correlation with the stretch. We compare their
        AIC, an estimator of the relative quality of statistical models that
    includes the number of free parameters, to determine which ones describe
best the data.}
    {Models with an evolution of the stretch with the redshift have a better
    AIC than the ones without.}
    {We find that implementing these models allows us to fit the data better
    than models without stretch evolution.}

@@

\section{Introduction}
Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
have enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate $z<0.5$, when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant (H$_0$), provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
the SNe~Ia are anchored on Cepheids \citep[the SH0ES
project][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
by cosmic microwave background (CMB) data measured by Planck assuming the
standard $\L$CDM\nn{,} or when the SN luminosity is anchored at intermediate
redshift by the baryon acoustic oscillation (BAO) scale
\citep{riess2019,reid2019,planck2016, feeney2019}.  While using the tip of the
red giant branch technique in place of the Cepheids seem to favor an
\mr{intermediate} value of H$_0$ \citep{freedman2019}, time delay measurements
from strong lensing seem to also favor high H$_0$ values \citep{wong2019}.

@@

For the last decades, numerous analysis have studied the relation between SNe~Ia
and host properties\nn{,} finding first that the standardized SNe~Ia magnitude
significantly depends on the host stellar-mass, SNe~Ia from high-mass host
\nn{being brighter on average}
\cite[e.g.][]{kelly2010, sullivan2010, childress2013, betoule2014, rigault2018, kim19}.
This mass-step correction is currently used in cosmological analyses
\citep[e.g.][]{betoule2014, scolnic2018a}, including for deriving H$_0$
\citep{riess2016, riess2019}. Yet, the underlying connection between the SN and
the host remains unclear when using global properties such as the host stellar
mass, which raises the question of the accuracy of the correction. More
recently, studies have used the local SN environment to probe more direct
connections between the SN and its environments \citep{rigault2013}\nn{,}
showing that local age-tracers such as the local specific star formation rate or
the local color are more strongly correlated with the standardized SN magnitude
\citep{rigault2018, roman2018, kim18}, suggesting the age to be the driving parameter
underlying the mass-step. If true, this would have a significant impact for
cosmology because the redshift drift magnitude correction to apply might
strongly vary \citep{rigault2013,childress2014,scolnic2018a}.  Yet, the
importance of local SN environmental study remains highly debated
\cite[e.g.][]{jones2015,jones2019} and especially the impact it has on the
derivation of H$_0$ \citep{jones2015,riess2016,riess2018,rose2019}.

In this paper, we take a step aside to probe the validity of our modeling of the
SN population that we claim to be constituted of two age-populations
\citep{rigault2013,rigault2015,rigault2018}: one old and one young, the former
having on average lower lightcurve stretches and being brighter after
standardization. We indeed use the correlation between the SN age probed with
local specific star formation rate in \cite{rigault2018} and the SN stretch to
model the expected evolution of the underlying SN stretch distribution as a
function of redshift. This modeling relies on three assumptions: (1) there
\nn{are indeed} 2 populations of SNe~Ia; (2) the relative fraction of each of
these populations as a function of redshift follows the age-modeling made in
\citep{rigault2018} and (3) the underlying distribution of stretch for each
age-sample can be modeled and do not significantly drift. This paper aims at
testing this with datasets from the literature.

% SERIES OF PAPER
%This paper is part of a series of papers where we test the validity of the two
%SNe Ia population models developed in \citep{rigault2018}. Briday et al. in
%prep will present how using different environmental tracer is affecting our
%ability to distinguish the two populations, and Rigault et al. in prep will
%further study the connection with H$_0$. Here we analyze predictions concerning
%the SNe Ia redshift drifts: (1) that the distribution of the SN~Ia stretch, a
%purely intrinsic SN~Ia property, is evolving as a function of redshift and (2)
%that this population-drift follows the relation that \citep{rigault2018}
%derived assuming that one population is “young“ with a rate proportional to the
%star formation rate in the Universe and the second is "old" with a rate
%proportional to the stellar mass.

The concept of the SNe~Ia age dichotomy aroused when the SN~Ia rate has first
been studied. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008}
have shown that the relative SNe~Ia rate in galaxies could only be explained if
two populations existed, one young, following the host star formation activity,
and one old following the host stellar mass (the so called “prompt and delayed“
or “A+B“ model). In \cite{rigault2018} we used the specific Star Formation Rate
at the SN location (Local sSFR or LsSFR) to classify which are the prompts
(those with a high LsSFR) and which are the delayed (those with low LsSFR).
Furthermore, since the first SNe~Ia host analysis, the SN stretch has been known
to be strongly correlated with the SN host properties \citep{hamuy1996,
hamuy2000}, which has been extensively confirmed since \citep[e.g.][]{neill2009,
sullivan2010, lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013,
rigault2013, pan2014, kim19}. Following the “A+B“ model and the connection between SN
stretch and host properties, \cite{howell2007} first discussed the potential
redshift drift of the SN stretch distribution. In this paper we revisit this
analysis with the most recent SNe~Ia dataset.

We present in section~\ref{sec:sample} the sample we are using for this
analysis, which is based on the Pantheon dataset \citep{scolnic2018a}. We
discuss the importance of obtaining a "complete" sample, i.e. representative of
the true underlying SNe Ia distribution, and how we build one from the Pantheon
sample.  We then present in section~\ref{sec:modeling} our modeling of the
distribution of stretch as a function of redshift based stretch distribution of
young and old SNe~Ia. Our results are presented in section~\ref{sec:results}
where we test if the SN stretch does evolve as a function of redshift and if the
age model is in good agreement with this evolution. Consequence for cosmology of
our results are briefly discussed in section~\ref{sec:bbc} and we conclude in
section~\ref{sec:ccl}.

%Throughout the paper we use the Planck 2015 cosmology from the astropy library.

\section{Complete Sample Construction}
\label{sec:sample}

We base our analysis on the latest SNe~Ia compilation: the Pantheon catalog from
\cite{scolnic2018a}. A naive approach to test the SN stretch redshift drift
would be to simply compare the observed SN stretch distributions in a few bins
of redshift.  In practice, selection effects are affecting the observed SN
stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
with the lightcurve stretch (and color), the first SN~Ia that a
magnitude-limited survey will miss are the lowest-stretch (and reddest) ones.
Consequently, if selection effects are not accounted for, one might confuse true
population drift with selection effects, and conversely \nn{\citep[e.g.,][]{kim19}}.

\nn{Assuming} sufficient (and random \yc{unbiased?}) spectroscopic follow-up for acquiring
typing and host redshift, the selection effects of magnitude-limited surveys
should be negligible bellow a given redshift where even the faintest \nn{SN~Ia}
can be observed. In contrast, targeted surveys have highly complex selection
functions and will be discarded from our analysis. Fortunately, modern SN
cosmology samples such as the Pantheon dataset are now dominated by magnitude
limited surveys.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
    \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$) lightcurve parameters
        of SNe~Ia from the SDSS, PS1 and SNLS samples from the Pantheon catalog.
        The individual SNe are shown as blue dots and a 2D histogram is
        shown in gray to highlight point density. The ellipse
        $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by the
        corresponding standardized absolute magnitude using the $\alpha$ and
        $\beta$ coefficients from \cite{scolnic2018a}. \nn{The lines represent
        the $(x_1, c)$ evolution for $m = m_{lim}$ in Eq.~\ref{eq:m_lim},
    for $z$ between $0.60$ and $z=1.00$ using SNLS's $m_{lim}$ of $24.8$~mag. The SNe
    above these lines are those for which $m > m_{lim}$, meaning they cannot be
    observed.}}
    \label{fig:maglim}
\end{figure}

We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color
distributions of SNe~Ia from \nn{the following surveys}: PanStarrs
\citep[PS1][]{rest2014}, the Sloan Digital Sky Survey
\citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plan with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates
the full distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011}
and \citet{campbell2013} for similar contours, the second using a more
conservative $|c| \leq 0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and
$c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$ along the aforementioned ellipse
given the standardization coefficient $\alpha=0.156$ and $\beta=3.14$ from
\cite{scolnic2018a}: the faintest SN~Ia is that with $(x_1=-1.66, c=0.25)$
and an absolute standardized magnitude at peak in Bessel-B band of
$M^{t_0}_{min} = -18.31$~mag. Since one need to detect this object typically a week
before and 10 days after peak to build a suitable lightcurve, the limiting
\nn{standardized} absolute magnitude effectively is approximately
$M_{min} = -18.00$~mag. Hence, given the \nn{\sout{$5\sigma$ point source detection}}
magnitude limit $m_{lim}$ of a magnitude limited survey, one can derive the
maximum redshift $z_{lim}$ above which the faintest SNe~Ia will be missed \nn{using the
relation between apparent magnitude, redshift and absolute magnitude given by:}
\begin{equation}\label{eq:m_lim}
    \nn{\cancel{m = \mu(z) + \underbrace{M_{min}^{t_0-10}(x_1, c)}_{=-18.00\;\text{mag}}
    < m_{lim}}}
\end{equation}
\begin{equation}\label{eq:m_lim}
    \yc{\mu(z_{lim}) = m_{lim} - M_{min}} \quad\text{min vs. lim???}
\end{equation}

SNLS typically acquires SNe~Ia in the redshift range $0.4<z<0.8$. At these
redshifts the rest-frame Bessel-B band roughly corresponds to the SNLS-$i$
filter that has a $5\sigma$ depth of 
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT final release website.}}. 
This converts to a $z_{lim}=0.60$, in perfect agreement with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. \nn{Fig.~14 of \citealt{perrett2010}
suggests a lower limit of $z_{lim}=0.55$; both will be considered as discussed below.}

@@

In similar redshift range, SDSS has a limiting magnitude of
22.5 \citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However,
the SDSS surveys were more sensitive to limited spectroscopic resources.
\cite{kessler2009} \nn{pointed out} that during year-1 of SDSS, SNe~Ia with
$r-mag<20.5$ were favored for spectroscopic follow up, corresponding to a
redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were used, and \cite{kessler2009} and
\cite{dilday2008} show a relative completeness up to $z_{lim}=0.2$. Following
these analyses we will use $z_{lim}=0.2$ as the baseline SDSS redshift limit \sout{for
the rest of \nn{this study}}.

For the rest of the analysis, we will consider \nn{\sout{both
cases,}} the \nn{aforementioned} redshift limits as well as more conservative cuts
for each of these surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for
PS1 and $z_{lim}=0.55$ for SNLS.

The complete sample selection is summarized in Table~\ref{tab:sample}, and the
redshift distribution of these three surveys are shown in Fig.~\ref{fig:cuts}.
We see there that the redshift limits we have selected roughly
correspond to the peak of these histograms. \nn{As the peaks indicate that we
are starting to miss the faintest SNe~Ia, it seems logical to have the redshift
cuts fall near said peaks.}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55.pdf}
    \caption{From top to bottom: histogram of SN~Ia redshifts for SDSS, PS1 and
    SNLS from the Pantheon dataset. The colored part of the
distribution shows which SNe~Ia has been kept in our analysis for they are
supposedly free from selection bias (see section~\ref{sec:sample}). The
dark (resp. light) color responds to the conservative (resp. fiducial) selection cut.}
    \label{fig:cuts}
\end{figure}

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory][]{aldering2004} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
4.2.2). \nn{Their spectroscopic follow-up were done on a redshift range of
$0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the searches were much
deeper. As such, these SNe~Ia} should also be a random sampling of the underlying
SN population. The SNf sample is particularly useful for studying SN property
drift as it enables us to have a large SN~Ia sample at $z<0.1$.

\nn{Finally, we included t}he HST sample
from Pantheon\nn{, that} follows the same logic of having a search deeper than the
follow-up and we therefore kept it entirely \citep{strolger04}, see
Table~\ref{tab:sample}. \nn{We present the stretch distribution and redshift histogram of these five surveys in Fig.~\ref{fig:sample}}.

\begin{table}
    \centering
    \caption{\nn{Source surveys and number of SNe~Ia used in our ``complete''
    sample. Conservative cuts are indicated in brackets.}}
    \label{tab:sample}
    \begin{tabular}{l c c}
    \hline\hline\\[-0.8em]
        Survey & z$_{lim}$ & N$_{\mathrm{SN}}$ \\[0.15em]
        \hline\\[-0.8em]
        SNf & -- & 114\\[0.30em]
        SDSS & 0.20(0.15) & 167(82)\\[0.30em]
        PS1 & 0.30(0.27) & 160(122)\\[0.30em]        
        SNLS & 0.60(0.55) & 102(78)\\[0.30em]
        HST & -- & 26\\[0.30em]
        \hline\\[-0.8em]
        Total & -- & 569(422)\\[0.30em]
        \hline
    \end{tabular}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75-lb.pdf}
    \caption{\nn{\textit{Main}: SN stretches in each survey as a function of redshift, using solid (resp. empty) markers for the
    conservative (resp. fiducial) selection. \textit{Top}: combined redshift 
    histogram for all surveys, in dark (resp. light) gray for the conservative (resp. fiducial) selection cuts.}}
    \label{fig:sample}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%
%   LITERATURE        %
%%%%%%%%%%%%%%%%%%%%%%%
%\mri{COMPARE HERE WITH LITERATURE. E.G. FIG. 6 OF SCOLNIC 2018, PS zmax
%AROUND0.27}

%\mri{Perrett et al. 2010 | SNLS | zmax ="z $\sim$ 0.6"}

%\mri{Dilday et al. 2008 | SDSS | sure good at z<0.12, Fig. 10 suggest still
%okat z=0.2}

%\mri{Kessler 2009 "For the SDSS-II, the image-subtraction
%pipeline efficiency (	subtr) is complete up to a redshift of z ∼ 0.2"}

%\mri{Kesslet 2009 | "For the SDSS-II and SNLS samples, the cutoff redshifts are
%estimated to be 0.15 and 0.65, respectively."}
% ADDIND: PS cut from Pantheon close to 0.27 see Fig. 6 of Scolnic 2018

>>>>

\begin{document}
\title{Redshift Evolution of the Underlying Type Ia Supernova Stretch
Distribution}

\author{N. Nicolas
\thanks{n.nicolas@ipnl.in2p3.fr, equal contribution}
\inst{1}
    \and M. Rigault
    \thanks{m.rigault@ipnl.in2p3.fr, equal contribution}
    \inst{2}
    \and Y. Copin\inst{1}   
    \and R. Graziani\inst{2}
    %HERE ALPHABETIC ORDER
    \and M. Briday\inst{1}
    \and Y.-L. Kim\inst{1}
    \and Saul Perlmutter\inst{3}
    \and G. Smadja\inst{1}   
}

\institute{Université de Lyon, F-69622, Lyon, France; Université de Lyon
    1, Villeurbanne; CNRS/IN2P3, Institut de Physique des Deux Infinis, Lyon
    \and 
    Université Clermont Auvergne, CNRS/IN2P3, Laboratoire de
    Physique de Clermont, F-63000 Clermont-Ferrand, France.
    \and
    Physics Division, Lawrence Berkeley National Laboratory, 
    1 Cyclotron Road, Berkeley, CA, 94720 
}

@@

\abstract{
{The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as
measure of survey statistic increase, the question of astrophysical systematic
uncertainties rises, and notably that of the SN~Ia population redshift drift.}
{In this paper, we study the redshift evolution of SN~Ia
\textsc{\texttt{SALT2.4}} lightcurve stretches, which is a purely intrinsic SN
property, to probe this drift.}
{The SN stretch has been shown to strongly correlate with the SN's environment,
and notably with stellar age tracers. Following the prediction of the fraction
of young and old SNe~Ia in nature from \cite{rigault2018}, and assuming
non-drifting underlying stretch distribution for each, we model the expected
SN~Ia stretch distribution as a function of redshift. We test our prediction
against data from the literature chosen to have negligible selection
effects to ensure that any observed drift is indeed astrophysical and not
artificial.}
{We clearly demonstrate that the underlying SNe~Ia stretch distribution is
evolving as a function of redshift, and that the young/old drifting model is a
much better description of the data than any non-drifting model, including the
sample-based asymmetric distributions used by the Beams with Bias Correction
algorithm, used to correct Malmquist bias. Our favored underlying stretch
model is bimodal: a high-stretch mode shared by both young and old environment,
and a low-stretch mode only existing in old environments.}
{The underlying SNe~Ia population is evolving as a function of redshift and the
exact impact on cosmology remains to be studied. Yet, the astrophysical drift of
the SN stretch distribution does affect current Malmquist bias corrections and
thereby distances derived from SN affected by selection effects. We highlight
that such a bias will increase with surveys covering larger redshift ranges,
which is particularly important for LSST.}

}

@@

\section{Introduction}
Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
have enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate $z<0.5$, when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant (H$_0$), provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
the SNe~Ia are anchored on Cepheids \citep[the SH0ES
project][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
by cosmic microwave background (CMB) data measured by Planck assuming the
standard $\L$CDM, or when the SN luminosity is anchored at intermediate
redshift by the baryon acoustic oscillation (BAO) scale
\citep{riess2019,reid2019,planck2016, feeney2019}. While using the tip of the
red giant branch technique in place of the Cepheids seem to favor an
intermediate value of H$_0$ \citep{freedman2019,freedman2020}, time delay
measurements from strong lensing seem to also favor high H$_0$ values
\citep{wong2019}.

@@

For the last decades, numerous analysis have studied the relation between SNe~Ia
and host properties, finding first that the standardized SNe~Ia magnitude
significantly depends on the host stellar-mass, SNe~Ia from high-mass host being
brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
betoule2014, rigault2018, kim19}.  This mass-step correction is currently used
in cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including
for deriving H$_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
between the SN and the host remains unclear when using global properties such as
the host stellar mass, which raises the question of the accuracy of the
correction. More recently, studies have used the local SN environment to probe
more direct connections between the SN and its environments \citep{rigault2013},
showing that local age-tracers such as the local specific star formation rate or
the local color are more strongly correlated with the standardized SN magnitude
\citep{rigault2018, roman2018, kim18}, suggesting age as the driving parameter
underlying the mass-step. If true, this would have a significant impact for
cosmology because the redshift drift magnitude correction to apply might
strongly vary \citep{rigault2013,childress2014,scolnic2018a}.  Yet, the
importance of local SN environmental study remains highly debated
\cite[e.g.][]{jones2015,jones2019} and especially the impact it has on the
derivation of H$_0$ \citep{jones2015,riess2016,riess2018,rose2019}.

In this paper, we take a step aside to probe the validity of our modeling of the
SN population which we claim to be constituted of two age-populations
\citep{rigault2013,rigault2015,rigault2018}: one old and one younger, the former
having on average lower lightcurve stretches and being brighter after
standardization. We use the correlation between the SN age probed with local
specific star formation rate in \cite{rigault2018} and the SN stretch to model
the expected evolution of the underlying SN stretch distribution as a function
of redshift. This modeling relies on three assumptions: (1) there are two
populations of SNe~Ia; (2) the relative fraction of each of these populations as
a function of redshift follows the age-modeling presented in \cite{rigault2018}
and (3) the underlying distribution of stretch for each age-sample can be
modeled and does not significantly drift. This paper aims at testing this with
datasets from the literature.

The concept of the SNe~Ia age dichotomy arose with study of the the SN~Ia rate.
\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have shown that
the relative SNe~Ia rate in galaxies could only be explained if two populations
existed, one young, following the host star formation activity, and one old
following the host stellar mass (the so called “prompt and delayed“ or “A+B“
model). In \cite{rigault2018} we used the specific Star Formation Rate at the SN
location (Local sSFR or LsSFR) to classify which are the younger (those with a
high LsSFR) and which are the older (those with low LsSFR). Furthermore, since
the first SNe~Ia host analyses, the SN stretch has been known to be strongly
correlated with the SN host properties \citep{hamuy1996, hamuy2000}, which has
been extensively confirmed since \citep[e.g.][]{neill2009, sullivan2010,
    lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013, rigault2013,
pan2014, kim19}. Following the “A+B“ model and the connection between SN stretch
and host properties, \cite{howell2007} first discussed the potential redshift
drift of the SN stretch distribution. In this paper we revisit this analysis
with the most recent SNe~Ia dataset.

We present in section~\ref{sec:sample} the sample we are using for this
analysis, which is based on the Pantheon dataset \citep{scolnic2018a}. We
discuss the importance of obtaining a ``complete'' sample, i.e. representative of
the true underlying SNe Ia distribution, and how we build one from the Pantheon
sample.  We then present in section~\ref{sec:modeling} our modeling of the
distribution of stretch as a function of redshift based models for the stretch
distribution of young and old SNe~Ia. Our results are presented in
section~\ref{sec:results}, where we test whether the SN stretch does evolve as a
function of redshift and if the age model is in good agreement with this
evolution. Consequences for cosmology of our results are briefly discussed in
section~\ref{sec:bbc} and we conclude in section~\ref{sec:ccl}.

\section{Complete Sample Construction}\label{sec:sample}

We base our analysis on the latest SNe~Ia compilation: the Pantheon catalog from
\cite{scolnic2018a}. A naive approach to test the SN stretch redshift drift
would be to simply compare the observed SN stretch distributions in a few bins
of redshift.  In practice, selection effects are affecting the observed SN
stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
with the lightcurve stretch (and color), the first SNe~Ia that a
magnitude-limited survey will miss are the lowest-stretch (and reddest) ones.
Consequently, if selection effects are not accounted for, one might confuse true
population drift with selection effects, and conversely.

Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring typing
and host redshift, the selection effects of magnitude-limited surveys should be
negligible below a given redshift where even the faintest SNe~Ia can be observed.
In contrast, targeted surveys have highly complex selection functions and will
be discarded from our analysis. Fortunately, modern SN cosmology samples such as
the Pantheon dataset are now dominated by magnitude limited surveys.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
    \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$)
        lightcurve parameters of SNe~Ia from the SDSS, PS1 and SNLS samples from
        the Pantheon catalog.  The individual SNe are shown as blue dots and a
        2D histogram is shown in gray to highlight point density. The ellipse
        $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by the corresponding
        standardized absolute magnitude using the $\alpha$ and $\beta$
        coefficients from \cite{scolnic2018a}. The grey diagonal lines represent
    the $(x_1, c)$ evolution for $m = m_{lim}$, for $z$ between $0.50$ and
$z=1.70$ using SNLS's $m_{lim}$ of $24.8$~mag.}
    \label{fig:maglim}
\end{figure}

We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color
distributions of SNe~Ia from the following surveys: PanStarrs
\citep[PS1][]{rest2014}, the Sloan Digital Sky Survey
\citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full distribution
\citep{guy2007,betoule2014}; see also \citet{bazin2011} and \citet{campbell2013}
for similar contours, the second using a more conservative $|c| \leq 0.2$ cut.
Assuming the SN absolute magnitude with $x_1=0$ and $c=0$ is $M_0=-19.36$
\citep{kessler2009,scolnic2014}, we can derive the absolute standardized
magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$ along the
aforementioned ellipse given the standardization coefficient $\alpha=0.156$ and
$\beta=3.14$ from \cite{scolnic2018a}: the faintest SN~Ia is that with
$(x_1=-1.66, c=0.25)$ and an absolute standardized magnitude at peak in Bessel-B
band of $M^{t_0}_{min} = -18.31$~mag. Since one ought to detect this object
typically a week before and 10 days after peak to build a suitable lightcurve,
the limiting standardized absolute magnitude effectively is approximately
$M_{lim} = -18.00$~mag. Hence, given the magnitude limit $m_{lim}$ of a
magnitude limited survey, one can derive the maximum redshift $z_{lim}$ above
which the faintest SNe~Ia will be missed using the relation between apparent
magnitude, redshift and absolute magnitude $\mu(z_{lim}) = m_{lim} - M_{lim}$.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55-cividis.pdf}
    \caption{\textit{From top to bottom}: Redshift histograms of SNe~Ia  from
        the SDSS, PS1 and SNLS dataset respectively (data from  Pantheon
        \citealt{scolnic2018a}).  The colored parts represent the distribution
        of SNe~Ia kept in our analysis for they are supposedly free from
        selection bias (see section~\ref{sec:sample}). The dark (resp. light)
    color responds to the conservative (resp. fiducial) selection cut.}
    \label{fig:cuts}
\end{figure}

SNLS typically acquires SNe~Ia in the redshift range $0.4<z<0.8$. At these
redshifts the rest-frame Bessel-B band roughly corresponds to the SNLS-$i$
filter that has a $5\sigma$ depth of
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{lim}=0.60$, in agreement with
\cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.  Fig.~14 of
\citealt{perrett2010} suggests a lower limit of $z_{lim}=0.55$; both will be
considered, as discussed below.

@@

In similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However, the
SDSS surveys were more sensitive to limited spectroscopic resources.
\cite{kessler2009} pointed out that during year-1 of SDSS, SNe~Ia with
$r-mag<20.5$ were favored for spectroscopic follow up, corresponding to a
redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were used, and \cite{kessler2009} and \cite{dilday2008}
show a relative completeness up to $z_{lim}=0.2$. Following these analyses we
will use $z_{lim}=0.2$ as the baseline SDSS redshift limit.

For the rest of the analysis, we will consider the aforementioned redshift
limits as well as more conservative cuts for each of these surveys, namely
$z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS1 and $z_{lim}=0.55$ for SNLS.

The complete sample selection is summarized in Table~\ref{tab:sample}, and the
redshift distribution of these three surveys are shown in Fig.~\ref{fig:cuts}.
As expected, the selected redshift limits roughly correspond to the peak of
these histograms.

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
4.2.2). Their spectroscopic follow-up was done over a redshift range of
$0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the searches were much
deeper. As such, these SNe~Ia should also be a random sampling of the underlying
SN population. The SNfactory sample is particularly useful for studying SN property
drift as it enables us to have a large SN~Ia sample at $z<0.1$.

Finally, we included the HST sample from Pantheon, that follows the same logic
of having a search deeper than the follow-up and we therefore kept it entirely
\citep{strolger04}, see Table~\ref{tab:sample}. We present the stretch
distribution and redshift histogram of these five surveys in
Fig.~\ref{fig:sample}.

\begin{table}
    \centering
    \caption{Composition of the SNe~Ia dataset used in this analysis.
    Conservative cuts are indicated in brackets.}
    \label{tab:sample}
    \begin{tabular}{l c c}
    \hline\hline\\[-0.8em]
        Survey & z$_{lim}$ & N$_{\mathrm{SN}}$ \\[0.15em]
        \hline\\[-0.8em]
        SNf & -- & 114\\[0.30em]
        SDSS & 0.20(0.15) & 167(82)\\[0.30em]
        PS1 & 0.30(0.27) & 160(122)\\[0.30em]        
        SNLS & 0.60(0.55) & 102(78)\\[0.30em]
        HST & -- & 26\\[0.30em]
        \hline\\[-0.8em]
        Total & -- & 569(422)\\[0.30em]
        \hline
    \end{tabular}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75-lb-cividis.pdf}
    \caption{\textit{Lower}: \textsc{\texttt{SALT2.4}} lightcurve stretch as a
        function of redshift (in log) for each survey considered in this
        analysis (see legend).  Full (resp. empty) markers correspond to the
        conservative (resp. fiducial) selection cuts. \textit{Top}: combined
        redshift histogram in dark (resp. light) gray for the conservative
    (resp. fiducial) selection cuts.}
    \label{fig:sample}
\end{figure}

\section{Modeling the redshift drift}\label{sec:modeling}

In~\cite{rigault2018} we presented a model for the evolution of the fraction of
younger and older SNe~Ia as a function of redshift following former work on
rates and delay time distributions \citep[e.g.,][]{mannucci2005,
scannapieco2005, sullivan2006, aubourg2008, childress2014, maozmannucci2014}.
In short, we assumed that the number of younger SNe~Ia follows the star
formation activity (SFR) in the Universe while the number of older SNe~Ia
follows the number of Gyr-old stars in the Universe, i.e. the stellar mass
(M$^*$). Hence, if we denote $\delta(z)$ (resp. $\psi(z) = 1-\delta(z)$) the
fraction of young (resp. old) SNe~Ia in the Universe as a function of redshift,
then their ratio $\delta/\psi$ is expected to follow the evolution of the
specific star formation rate (SFR/M$^*$) which goes as $(1+z)^{2.8}$ until
$z\sim2$ \citep[e.g.,][]{tasca2015}. Since $\delta(0.05) \sim \psi(0.05)$
\citep{rigault2013,rigault2018,wiseman2020}, in agreement with rate expectations
\citep{mannucci2006,rodney2014}, we concluded in \cite{rigault2018} that
\begin{equation}
    \label{eq:delta}
    \delta(z) = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}
\end{equation}
with $K=0.87$. This model is comparable to the evolution predicted by
\cite{childress2014} based on SN rates in galaxies depending on their quenching
time as a function of their stellar mass.

\subsection{“Base” underlying stretch distribution}
\label{sec:basemodel}
